Note on Boundary condition and choice of grid

### 2024-07-08 write2/manuscript_part2_2024.tex
worked on part 4

%alpha
The surface veering $\alpha_\star$ is key to quantify the Ekman pumping and of central relevance for larger-scale applications. For the LES, we use the first derivatives of the velocity components calculated by 3$^\text{rd}$-order forward finite differences to estimate $\alpha_\star$:
\begin{align}
  \alpha_\star = \arctan\left(\left.\left\langle \frac{\partial v}{\partial z}\right\rangle_0\right/\left\langle \frac{\partial u}{\partial z}\right\rangle_0\right)-\arctan\left(\frac{V_G}{U_G}\right).
\end{align}
 Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$. In general, we observe an increase of $\alpha_\star$ with resolution. The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 still 0.9° is missing. Fig. \ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little. In contrast to the higher $\RE$, the veering of the wind vector continues in the lower parts of the boundary layer. Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point. In general, the finer resolved simulations show veering angles that are very close to the theoretical prediction.

@@

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_profiles_outer.png}
	}
  \caption{Shear-aligned velocity profiles in outer scaling. Solid lines: u-component, dashed lines: v-component. The lowest grid point is marked by $x$}
  \label{velocity_outer}
\end{figure}

% profiles
Figure \ref{velocity_outer} shows the velocity profiles from the LES and the theoretical profiles for each of the $\RE$. In general, the velocity profiles of the LES agree well with the theoretical profiles: The course of the theoretical Ekman layer is matched and the simulations exhibit a logarithmic layer for the U-component. The lowest points of the LES of the higher $\RE$ fall into the logarithmic layer. For Re1, the lowest grid point falls into the buffer layer, which is visible as the curved course of the U-component below the logarithmic layer in the theoretical profile. The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid walls \cite{brasseur2010designing}. This log-layer mismatch arises from a competition between the scales $u_\star$ and $z$ and other velocity and length scales introduced by the discretization of the dynamical system \citep{mason1992stochastic,brasseur2010designing}. In other words, at the lower boundary, the relevant eddies are too small to be resolved by the grid and their contribution to the flow has to be modeled. Also the vertical component is massively restricted by the non-permeability of the wall, known as blocking effect. On the contrary, the SGS-closure assumes isotropic turbulence, which is not the case on the grid scale near the wall.

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_kappa}
}
  \caption{K\'arm\'an measure $\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
  \label{3Re_kappa}
\end{figure}

% kappa
In the logarithmic region, the profile of the u-component should follow the logarithmic law of eq. \ref{eqn:log-law} with the K\'arm\'an measure $\kappa = (z^+\partial U^+/\partial z^+)^{-1}=const.$. In the low-Reynolds-number case, the viscous sublayer represents about $0.5\%$ of the boundary layer, while this layer is not visible for the high-Reynolds-number cases, where the share drops to $~10^{-5}-10^{-7}$. Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3. In order to estimate $\kappa_{LES}$ from the simulations, we perform a linear regression of the mean velocity in x-direction and the logarithm of height between the seventh grid point and the grid point corresponding to the height $z^-=0.1$ for $Re1\_X$ and $z^-=0.12$ for $Re2\_X$ and $Re3\_X$. We consider only points above the 7th grid point following arguments of \cite{maronga2020improved}, that in PALM, the mean velocity profiles follow MOST above the seventh grid point. The number of values for each regression is 6, 12, and 18 for $ReX\_100$, $ReX\_150$ and $ReX\_200$, respectively.  We do not compute $\kappa$ for $ReX\_50$ since there are only six grid points below $z^-\approx 0.12$. The resulting $\kappa_{LES}$ for the other simulations are shown in Tbl. \ref{simulation_results}. The coefficient of determination is above 0.99 for all fits. $\kappa_{LES}$ is decreasing with finer resolution for all $\RE$. While the values for $ReX\_100$ are rather high with a value around $0.53$, finer resolution leads to values closer to measurements around $0.46$ for $Re1\_200$ and around $0.44$ for $Re2\_200$ and $Re3\_200$. The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in $Re1\_dyn$ we even see $\kappa_{LES}=0.39$, while $Re2\_dyn$ and $Re3\_dyn$ yield $0.42$ and $0.43$, respectively. Figure \ref{3Re_kappa} shows the K\'arm\'an measure in the logarithmic layer. For the simulations $ReX\_50$, no constant $\kappa$ can be observed. An increase in resolution leads to a profile of the K\'arm\'an measure closer to the theoretical curve for all cases. Close to the bottom, the K\'arm\'an measure is heavily influenced by the proximity of the wall and rather a function of the vertical index than of the physical distance from the wall. In accordance with the observations of \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations. Above, the curve smoothens and approaches the expected value of $\kappa$ (at least for the finer resolutions).

@@

As expected for an LES with isotropic grid spacing, the resolution near the surface is a critical part of the simulation. We can support our above observations by taking a look at the two-point correlation in the critical layers of the simulations. The two point correlation is defined as 
\begin{align}
  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2.
\end{align}
As suggested by \cite{wurps2020grid}, in an isotropic grid with periodic horizontal boundary conditions the turbulent structures of the velocity's w-component in y-direction can serve as an indicator how well the flow is resolved. Figure \ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer. We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure). The number of cells by which the average structure is resolved is equal to $\sigma/\Delta$. All $\RE$ show a similar behaviour: the average size of the structure decreases with finer resolution and seems to approach a limit. The structure size of ReX\_150 and ReX\_200 are very close together, which indicates beginning convergence. Furthermore, the number of cells by which the structures are resolved exceed 4 from the resolution of ReX\_150 on. This corresponds to the resolution from which on a logarithmic layer with a constant $\kappa$ can be seen. It also fits to the rule of thumb given in \cite{wurps2020grid}, that in a sufficiently resolved part of an LES the average structures of the w-component in y-direction should be resolved by at least 4 cells. The dynamic closure shows structures that are slightly smaller than the structures of $ReX\_200$, which is caused by a tendency to lower eddy-viscosities and, hence, a weaker coupling of neighboring grid cells.

>>>>

%alpha
The surface veering $\alpha_\star$ is key to quantify the Ekman pumping and of central relevance for larger-scale applications. For the LES, we use the first derivatives of the velocity components calculated by 3$^\text{rd}$-order forward finite differences to estimate $\alpha_\star$:
\begin{align}
  \alpha_\star = \arctan\left(\left.\left\langle \frac{\partial v}{\partial z}\right\rangle_0\right/\left\langle \frac{\partial u}{\partial z}\right\rangle_0\right)-\arctan\left(\frac{V_G}{U_G}\right).
\end{align}
Nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$ (cf. Tab.~\ref{simulation_results}), where we observe a slight increase of $\alpha_\star$ with resolution; The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 the offset is 0.9°. The hodograph (Fig.~\ref{3Re_hodograph}) illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 closely follow the theoretical hodograph; in fact, the turning angle at the respective height is actually slightly overestimated. Rather, the offset is due to the veering of the wind vector continues in the lower part of the boundary layer that is not present in the LES. Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point. In general, the finer resolved simulations show veering angles that are very close to the theoretical prediction.

@@

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_profiles_outer.png}
	}
  \caption{Shear-aligned velocity profiles in outer scaling. Solid lines: u-component, dashed lines: v-component. The lowest grid point is marked by a cross ($\times$) colored according to the corresponding resolution.}
  \label{velocity_outer}
\end{figure}

% profiles
Velocity profiles from LES and our theoretical framework agree well for the different $\RE$ in Fig.\,\ref{velocity_outer}: Most prominently, the rotation across the Ekman layer agrees and a logarithmic layer for the U-component is present. The lowest points of the LES of the higher $\RE$ fall into the logarithmic layer. For Re1, the lowest grid point falls into the buffer layer, which is visible as the curved course of the U-component below the logarithmic layer in the theoretical profile. The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid wall \citep{brasseur2010designing}. This log-layer mismatch arises from a competition between the scales $u_\star$ and $z$ and other velocity and length scales introduced by the discretization of the dynamical system \citep{mason1992stochastic,brasseur2010designing}. In other words, at the lower boundary, the relevant eddies are too small to be resolved by the grid and their contribution to the flow has to be modeled. Also the vertical component is restricted by the wall's impermeability, known as blocking effect-- both effects illustrate a violation of isotropy near the wall and thus counter the SGS-closure assumptions.

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_kappa}
}
  \caption{K\'arm\'an measure $\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
  \label{fig:3Re_kappa}
\end{figure}

% kappa
The goodness-of-fit in the logarithmic layer (Eq.\,\ref{eqn:log-law}) is assessed via the K\'arm\'an measure~(Fig.\ref{fig:3Re_kappa})
\begin{align}
  \kappa(z)=(z^+\partial U^+/\partial z^+)^{-1}\simeq const.
\end{align}
which is estimated here by linear regression of the mean velocity in x-direction against the logarithm of height between the seventh grid point and the $z^-=0.1$ for Re1 and $z^-=0.12$ for Re2 and Re3. The seventh grid point is chosen as lower boundary for the logarithmic layer as, in PALM, the mean velocity profile follows MOST aloft \citep{maronga2020improved}. The number of values for each regression is 6, 12, and 18 for ReX\_100, ReX\_150 and ReX\_200, respectively.  In case Re1, the viscous sublayer occupies about $0.5\%$ of the boundary layer, but this share drops to $~10^{-5}-10^{-7}$ for cases Re2 and Re3 respectively. Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3. We do not compute $\kappa$ for ReX\_50 since there are only six grid points below $z^-\approx 0.12$, and o constant K\'arm\'an measure $\kappa(z)$ can be observed. The resulting best estimates for $\kappa_{LES}$ are shown in Tab.\,\ref{simulation_results}, where the coefficient of determination ('r-squared') is above 0.99 for all cases.

For finer resolution, $\kappa_{LES}$ decreases with increasing Re for the range of scale separation investigated here. ReX\_100 yields a value around $0.53$ (which is outside the range agreed on by atmospheric observation and laboratory measurements), the finer resolution yields values around $0.46$ for Re1\_200 and around $0.44$ for Re2\_200 and Re3\_200, which is in excellenet agreement with DNS data and laboratory measurements. The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in Re1\_dyn we even see $\kappa_{LES}=0.39$, while Re2\_dyn and Re3\_dyn yield $0.42$ and $0.43$, respectively. Close to the bottom, the K\'arm\'an measure is heavily influenced by the proximity of the wall and rather a function of the vertical index than of the physical distance from the wall. In accordance with \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations. Above, the curve smoothens and approaches the expected value of $\kappa$.

@@

The resolution near the surface is critical for an LES with isotropic grid spacing. \todo{Missing transition from resolution to two-point correlation!} The two-point correlation is defined as 
\begin{align}
  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2,
\end{align}
and it supports the above findings \todo{which findings?}. 
For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the velocity's w-component in the y-direction indicate how well the flow is resolved \citep{wurps2020grid}. Fig.\,\ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer. We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure). The number of cells by which the average structure is resolved is equal to $\sigma/\Delta$. All $\RE$ show a similar behaviour: the average size of the structure decreases with finer resolution and seems to approach a limit. The structure size of ReX\_150 and ReX\_200 are very close together, which indicates beginning convergence. Furthermore, the number of cells by which the structures are resolved exceed 4 from the resolution of ReX\_150 on. This corresponds to the resolution from which on a logarithmic layer with a constant $\kappa$ can be seen. It also fits to the rule of thumb given in \cite{wurps2020grid}, that in a sufficiently resolved part of an LES the average structures of the w-component in y-direction should be resolved by at least 4 cells. The dynamic closure shows structures that are slightly smaller than the structures of $ReX\_200$, which is caused by a tendency to lower eddy-viscosities and, hence, a weaker coupling of neighboring grid cells.
